Building socket-aware BPF programs

### 2018-10-30 paper/paper.tex
paper: Improve stateful firewalling section
paper: Write up a second use case

This paper is laid out as follows. Firstly, we discuss the constraints that the
BPF verifier places upon BPF programs and how those constraints translate into
requirements on the implementation of socket lookup helpers in the Linux BPF
API. The implementation of reference tracking functionality is presented.
Secondly, we describe extensions to the BPF API to provide socket introspection
capabilities, taking into account flexibility and performance. Thirdly, we
present some use cases for the new functionality, including more detail on the
stateful firewall case and a description of the use of the new functionality
for load-balancing. Finally, we conclude and assess the position of this work.

@@

\todo{Associate custom metadata with socket?}

@@

Stateful firewalling is distinguished from stateless firewalling in that it
provides the ability to associate two directions of a connection together and
apply a filtering policy based upon the direction of the connection. This
commonly involves using a connection tracker to reconstruct the connection
state of each endpoint of the connection without specific knowledge from those
endpoints~\cite{conntrack,ovs-ct}. In cases where Linux is configured as a
firewall between applications that exist in distinct Linux kernel copies
running in the network, this makes sense; The information that is needed to
perform stateful firewalling is not present on the Linux instance that needs
this information, so the information must either be shared from the peers
(assuming that the firewall orchestrator has control over the peer endpoints),
or this information must be synthesised based upon the packets that are seen on
this intermediate Linux instance. However, in many cases this model is used
when filtering traffic where one endpoint is co-located with the Linux stack
that performs the firewalling, which already contains information about the
local end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that Linux already creates and maintains.

Using the socket lookup functions described in this paper, in conjunction with
the socket properties including local Layer~3 and Layer~4 information available
in the \verb+bpf_sock+ (Listing~\ref{lst:helper-struct}), BPF programs can
identify the directionality of connections for packets ingressing or egressing
the Linux stack on a network device.

\subsection{Load-balancing}

\todo{Early in the stack, figure out whether the traffic is destined for a
local process and if so, pass. Otherwise, perform load-balancing to another
node where the traffic is destined for.}

@@

>>>>

This paper is laid out as follows. Firstly, we discuss the constraints that the
BPF verifier places upon BPF programs and how those constraints translate into
requirements on the implementation of socket lookup helpers in the Linux BPF
API. The implementation of reference tracking functionality is presented.
Secondly, we describe extensions to the BPF API to provide socket introspection
capabilities, taking into account flexibility and performance. Thirdly, we
present some use cases for the new functionality, including more detail on the
stateful firewall case and an additional usage in handling management traffic
in forwarding devices. Finally, we conclude and assess the position of this
work.

@@

%\todo{Associate custom metadata with socket?}

@@

Stateful firewalling is distinguished from stateless firewalling in that it
provides the ability to associate two directions of a connection together and
apply a filtering policy based upon the direction of the connection. This
commonly involves using a connection tracker to reconstruct the connection
state of each endpoint of the connection without specific knowledge from those
endpoints~\cite{conntrack,ovs-ct}.

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this makes sense;
The information that is needed to perform stateful firewalling is not present
on the Linux instance that needs this information, so the information must
either be shared from the peers (assuming that the firewall orchestrator has
control over the peer endpoints), or this information must be synthesised based
upon the packets that are seen on this intermediate Linux instance. However, in
many cases this model is used when filtering traffic where one endpoint is
co-located with the Linux stack that performs the firewalling, which already
contains information about the local end of the connection. In this use case,
the connection tracking table effectively duplicates knowledge that Linux
already creates and maintains.

Using the socket lookup functions described in this paper, in conjunction with
the socket properties including local Layer~3 and Layer~4 information available
in the \verb+bpf_sock+ (Listing~\ref{lst:helper-struct}), BPF programs can
identify the directionality of connections for packets ingressing or egressing
the Linux stack on a network device. Between this directionality information
and a network policy provided by a userspace program, BPF programs are able to
identify locally-sourced or locally-destined traffic and allow or deny traffic
ingressing or egressing the Linux stack.

\subsection{Filtering management traffic in forwarding devices}

A common design of devices providing routing and forwarding functionality is to
place high-bandwidth forwarding elements into a system with a low-bandwidth
CPU. In many cases, the goal of the CPU is to only to handle control traffic
for configuring the forwarding path, which is a fraction of the traffic handled
by the actual forwarding path. Existing implementations may direct only traffic
for a specific IP into the management CPU, or could plausibly use XDP to
configure DoS filters so that only legitimate traffic is forwarded up the
stack. Such implementations may not provide the degree of restriction on
locally destined traffic that is desired, or may require additional
co-ordination with the socket layer that requires significant orchestration.

Using the upcoming XDP support for socket lookup, early in the driver level of
the Linux stack the traffic could be compared with the socket table and only
passed up the stack if the traffic corresponds to a local application. As this
functionality uses the socket table directly, it is always kept in sync with
the current applications without additional orchestration of BPF maps. Traffic
which is not intended to be handled locally can then be either handled by
additional BPF logic programmed in the XDP hook.

@@

%\todo{Associate custom metadata with socket?}
